---
title: "Advanced Features in cnma"
author: "Mahmood Developer"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Advanced Features in cnma}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

## Introduction

This vignette demonstrates advanced features of the **cnma** package, including:

- Transportability analysis
- GRADE quality weighting
- Meta-regression
- Publication bias assessment
- Leave-one-out analysis

## Setup

```{r library, eval = FALSE}
library(cnma)

# Generate rich example data
data <- simulate_cnma_data(n_studies = 40, seed = 123)
```

## Transportability Analysis

Transportability analysis weights studies by their similarity to a target population.

### Define Target Population

```{r target, eval = FALSE}
# Define characteristics of your target population
target_population <- data.frame(
  age_mean = 68,
  female_pct = 0.52,
  bmi_mean = 27.5,
  charlson = 1.8
)
```

### Configure Transportability

```{r transport-config, eval = FALSE}
config <- setup_cnma(
  use_transport = TRUE,
  transport_metric = "mahalanobis",  # Distance metric
  transport_kernel = "gaussian",      # Weighting kernel
  transport_truncation = 0.02,        # Truncate extreme weights
  use_bayesian = FALSE,
  export_plots = FALSE
)
```

### Run Analysis

```{r transport-run, eval = FALSE}
results <- run_cnma_analysis(
  data = data,
  target_population = target_population,
  config = config
)
```

### Distance Metrics

Available metrics:
- `"mahalanobis"`: Accounts for correlations between covariates (recommended)
- `"euclidean"`: Simple Euclidean distance

### Kernel Functions

Available kernels:
- `"gaussian"`: Smooth decay of weights
- `"tricube"`: Sharper cutoff for distant studies

## GRADE Quality Weighting

Weight studies by evidence quality using GRADE assessments.

```{r grade, eval = FALSE}
# Ensure your data has a 'grade' column
table(data$grade)

config <- setup_cnma(
  use_grade_weighting = TRUE,
  use_bayesian = FALSE,
  export_plots = FALSE
)

results <- run_cnma_analysis(data, config = config)
```

Quality weights (default):
- High: 1.0
- Moderate: 0.75
- Low: 0.5
- Very low: 0.25

## Risk of Bias (RoB2) Weighting

```{r rob2, eval = FALSE}
# Ensure your data has a 'rob2' column
table(data$rob2)

config <- setup_cnma(
  use_rob2_weighting = TRUE,
  rob2_var = "rob2",
  use_bayesian = FALSE
)

results <- run_cnma_analysis(data, config = config)
```

Bias weights (default):
- Low risk: 1.0
- Some concerns: 0.75
- High risk: 0.5

## Study Design Weighting

```{r design, eval = FALSE}
config <- setup_cnma(
  use_design_weighting = TRUE,
  design_var = "is_rct",
  design_weight_map = c(`1` = 1.0, `0` = 0.6),  # RCT vs observational
  use_bayesian = FALSE
)

results <- run_cnma_analysis(data, config = config)
```

## Meta-Regression

Investigate effect modifiers with meta-regression.

### Basic Meta-Regression

```{r metareg, eval = FALSE}
config <- setup_cnma(
  run_metareg = TRUE,
  metareg_covariates = c("age_mean", "female_pct", "bmi_mean"),
  use_bayesian = FALSE
)

results <- run_cnma_analysis(data, config = config)
```

### Spline Transformations

For non-linear relationships:

```{r spline, eval = FALSE}
config <- setup_cnma(
  run_metareg = TRUE,
  metareg_covariates = c("age_mean", "female_pct", "bmi_mean"),
  metareg_spline_covars = c("age_mean", "bmi_mean"),
  metareg_spline_df = 3,
  metareg_spline_cv = TRUE,  # Cross-validation for df selection
  metareg_cv_folds = 5,
  use_bayesian = FALSE
)

results <- run_cnma_analysis(data, config = config)
```

### Cluster-Robust Variance

For studies with shared characteristics:

```{r cr2, eval = FALSE}
config <- setup_cnma(
  run_metareg = TRUE,
  metareg_covariates = c("age_mean"),
  metareg_cr2 = TRUE,  # CR2 variance correction
  use_bayesian = FALSE
)

results <- run_cnma_analysis(data, config = config)
```

## Publication Bias Assessment

### Enable Publication Bias Tests

```{r pub-bias, eval = FALSE}
config <- setup_cnma(
  enable_egger = TRUE,        # Network-level Egger test
  run_trimfill = TRUE,        # Trim-and-fill analysis
  run_selection_models = TRUE,
  run_copas = TRUE,
  pet_peese_min_k = 10,       # Minimum studies for PET-PEESE
  use_bayesian = FALSE
)

results <- run_cnma_analysis(data, config = config)
```

## Leave-One-Out Analysis

Assess robustness by removing studies one at a time.

```{r loo, eval = FALSE}
config <- setup_cnma(
  run_loo = TRUE,              # Leave-one-out
  run_loto = TRUE,             # Leave-one-treatment-out
  use_bayesian = FALSE
)

results <- run_cnma_analysis(data, config = config)
```

## Inconsistency Assessment

### Node-Splitting

```{r nodesplit, eval = FALSE}
config <- setup_cnma(
  nodesplit_min_k = 10,
  use_bayesian = FALSE
)

results <- run_cnma_analysis(data, config = config)
```

### UME Inconsistency Model

```{r ume, eval = FALSE}
config <- setup_cnma(
  enable_ume = TRUE,
  use_bayesian = FALSE
)

results <- run_cnma_analysis(data, config = config)
```

### Net Heat Plot

```{r netheat, eval = FALSE}
config <- setup_cnma(
  run_netheat = TRUE,
  use_bayesian = FALSE
)

results <- run_cnma_analysis(data, config = config)
```

## Bayesian Analysis

For more complex models and probabilistic inference:

```{r bayesian, eval = FALSE}
config <- setup_cnma(
  use_bayesian = TRUE,
  bayes_link = "identity",
  bayes_likelihood = "normal",
  bayes_chains = 3,
  bayes_iter = 10000,
  bayes_warmup = 5000,
  bayes_nodesplit = TRUE
)

# Note: Requires JAGS installation
results <- run_cnma_analysis(data, config = config)
```

### Bayesian Meta-Regression

```{r bayes-metareg, eval = FALSE}
config <- setup_cnma(
  use_bayesian = TRUE,
  bayes_metareg_covariates = c("age_mean", "female_pct"),
  bayes_chains = 3,
  bayes_iter = 10000
)

results <- run_cnma_analysis(data, config = config)
```

## SUCRA Analysis

Surface Under the Cumulative Ranking curve.

### Basic SUCRA

```{r sucra, eval = FALSE}
# SUCRA is calculated automatically with main analysis
# Access via results$results$main_nma
```

### Bootstrap SUCRA

For uncertainty quantification:

```{r sucra-boot, eval = FALSE}
config <- setup_cnma(
  sucra_boot = TRUE,
  sucra_boot_iter = 300,
  use_bayesian = FALSE
)

results <- run_cnma_analysis(data, config = config)
```

## Combining Features

You can combine multiple advanced features:

```{r combined, eval = FALSE}
# Comprehensive analysis
config <- setup_cnma(
  # Weighting
  use_transport = TRUE,
  use_grade_weighting = TRUE,
  use_rob2_weighting = TRUE,

  # Meta-regression
  run_metareg = TRUE,
  metareg_covariates = c("age_mean", "female_pct", "bmi_mean"),
  metareg_spline_covars = c("age_mean"),

  # Diagnostics
  run_loo = TRUE,
  run_loto = TRUE,
  enable_egger = TRUE,
  run_netheat = TRUE,

  # Output
  export_plots = TRUE,
  plot_dir = "comprehensive_plots",
  report_html = TRUE,

  # Performance
  parallel_strategy = "multicore",
  n_cores = 4,

  use_bayesian = FALSE
)

target <- data.frame(
  age_mean = 65,
  female_pct = 0.5,
  bmi_mean = 28,
  charlson = 1.5
)

results <- run_cnma_analysis(
  data = data,
  target_population = target,
  config = config
)
```

## Exporting Results

### Save Plots

```{r export-plots, eval = FALSE}
config <- setup_cnma(
  export_plots = TRUE,
  plot_dir = "my_nma_plots"
)
```

### Generate HTML Report

```{r html-report, eval = FALSE}
config <- setup_cnma(
  report_html = TRUE,
  report_file = "nma_report.html"
)
```

### Write Manifest

```{r manifest, eval = FALSE}
config <- setup_cnma(
  write_manifest = TRUE
)
```

## Performance Optimization

### Parallel Processing

```{r parallel, eval = FALSE}
config <- setup_cnma(
  parallel_strategy = "multicore",
  n_cores = 4
)

# Or enable manually
cnma_parallel_on("multicore", workers = 4)
results <- run_cnma_analysis(data, config = config)
cnma_parallel_off()
```

### Caching

```{r cache, eval = FALSE}
config <- setup_cnma(
  enable_cache = TRUE
)
```

### Logging

```{r logging, eval = FALSE}
config <- setup_cnma(
  log_to_file = TRUE,
  log_file = "cnma_analysis.log"
)
```

## Tips and Best Practices

1. **Transportability**: Use when applying results to a different population
2. **GRADE Weighting**: Use when you have quality assessments for all studies
3. **Meta-regression**: Pre-specify covariates; beware of overfitting
4. **Publication Bias**: Only meaningful with sufficient studies (>10)
5. **Bayesian Analysis**: Provides probabilistic statements; requires JAGS
6. **Parallel Processing**: Benefits diminish with small networks (<20 studies)

## References

- Guyatt GH et al. GRADE guidelines. *Journal of Clinical Epidemiology* 2011.
- Sterne JAC et al. RoB 2: a revised tool for assessing risk of bias in randomised trials. *BMJ* 2019.
- Dias S et al. Network meta-analysis for decision-making. Wiley, 2018.
