---
title: "Getting Started with cnma"
author: "Mahmood Developer"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with cnma}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

## Introduction

The **cnma** package provides a comprehensive toolkit for conducting network meta-analysis (NMA) with both frequentist and Bayesian approaches. This vignette will guide you through the basic workflow.

## Installation

```{r, eval = FALSE}
# Install from GitHub
remotes::install_github("mahmood726-cyber/HFN786")
```

## Quick Start

### 1. Load the Package

```{r library}
library(cnma)
```

### 2. Generate Example Data

For demonstration purposes, we'll use simulated data:

```{r simulate}
# Generate network with 30 studies and 5 treatments
data <- simulate_cnma_data(n_studies = 30, seed = 42)

# Examine the data structure
head(data)
```

The data contains:
- `studlab`: Study identifier
- `treat1`, `treat2`: Treatment comparison
- `TE`: Treatment effect (log hazard ratio)
- `seTE`: Standard error
- Additional covariates: `age_mean`, `female_pct`, `bmi_mean`, etc.

### 3. Configure Analysis

Create a configuration object with your analysis preferences:

```{r config}
config <- setup_cnma(
  sm = "HR",              # Summary measure: Hazard Ratio
  use_bayesian = FALSE,   # Use frequentist approach
  use_transport = FALSE,  # Disable transportability for quick demo
  use_grade_weighting = FALSE,
  run_metareg = FALSE,
  export_plots = FALSE,
  report_html = FALSE
)
```

### 4. Run Analysis

```{r analysis, eval = FALSE}
# Run the main analysis
results <- run_cnma_analysis(
  data = data,
  ref_treatment = "Placebo",
  config = config
)

# View results
print(results)
summary(results)
```

## Data Format

Your data must contain these required columns:

- `studlab`: Study identifier (character)
- `treat1`: First treatment in comparison (character)
- `treat2`: Second treatment in comparison (character)
- `TE`: Treatment effect estimate (numeric, log-scale for ratios)
- `seTE`: Standard error of treatment effect (numeric, > 0)

Optional columns for advanced features:

- `grade`: GRADE quality ("High", "Moderate", "Low", "Very low")
- `is_rct`: Study design (1 = RCT, 0 = observational)
- `rob2`: Risk of bias ("low", "some concerns", "high")
- Covariates: `age_mean`, `female_pct`, `bmi_mean`, `charlson`, etc.

## Network Characteristics

```{r network, eval = FALSE}
# Number of studies
n_studies <- length(unique(data$studlab))
cat("Number of studies:", n_studies, "\n")

# Number of treatments
treatments <- unique(c(data$treat1, data$treat2))
cat("Number of treatments:", length(treatments), "\n")
cat("Treatments:", paste(treatments, collapse = ", "), "\n")

# Number of comparisons
cat("Number of comparisons:", nrow(data), "\n")
```

## Configuration Options

The `setup_cnma()` function provides extensive configuration options:

### Summary Measure
```{r sm-options, eval = FALSE}
config <- setup_cnma(sm = "OR")  # Odds Ratio
config <- setup_cnma(sm = "RR")  # Risk Ratio
config <- setup_cnma(sm = "HR")  # Hazard Ratio
config <- setup_cnma(sm = "MD")  # Mean Difference
```

### Analysis Methods
```{r methods, eval = FALSE}
# Frequentist only (fast)
config <- setup_cnma(use_bayesian = FALSE)

# Add Bayesian analysis (slower, requires JAGS)
config <- setup_cnma(
  use_bayesian = TRUE,
  bayes_chains = 3,
  bayes_iter = 10000
)
```

### Advanced Features
```{r advanced, eval = FALSE}
# Transportability analysis
config <- setup_cnma(
  use_transport = TRUE,
  transport_metric = "mahalanobis"
)

# Meta-regression
config <- setup_cnma(
  run_metareg = TRUE,
  metareg_covariates = c("age_mean", "female_pct", "bmi_mean")
)

# Quality weighting
config <- setup_cnma(
  use_grade_weighting = TRUE,
  use_rob2_weighting = TRUE
)
```

## Understanding Results

The results object contains:

- `results$main_nma`: Main network meta-analysis (netmeta object)
- `results$data`: Input data
- `config`: Configuration used
- `ref_treatment`: Reference treatment

### Heterogeneity Statistics

```{r heterogeneity, eval = FALSE}
# Tau-squared (between-study variance)
results$results$main_nma$tau

# I-squared (percentage of variation due to heterogeneity)
results$results$main_nma$I2.random * 100
```

## Validation

Before running analysis, validate your data:

```{r validation, eval = FALSE}
# Check for required columns and valid values
validate_cnma_input(data)

# Clean data (remove invalid rows)
clean_data <- cnma_clean_data(data)
```

## Parallel Processing

For large networks, enable parallel processing:

```{r parallel, eval = FALSE}
# Enable multicore processing
cnma_parallel_on("multicore", workers = 4)

# Run analysis
results <- run_cnma_analysis(data, config = config)

# Disable when done
cnma_parallel_off()
```

## Next Steps

- See `vignette("advanced-features")` for meta-regression and transportability
- See `vignette("bayesian-analysis")` for Bayesian methods
- See `?run_cnma_analysis` for complete function documentation

## References

- Rucker G, Schwarzer G. Ranking treatments in frequentist network meta-analysis works without resampling methods. *BMC Medical Research Methodology* 2015; 15:58.
- Dias S, Welton NJ, Caldwell DM, Ades AE. Checking consistency in mixed treatment comparison meta-analysis. *Statistics in Medicine* 2010; 29(7-8):932-944.
