---
title: "Publication-Ready Network Meta-Analysis with CNMA"
author: "CNMA Development Team"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Publication-Ready Network Meta-Analysis with CNMA}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8,
  fig.height = 6
)
```

## Introduction

This vignette demonstrates how to conduct a publication-ready network meta-analysis (NMA) using the **cnma** package, following best practices from leading statistics journals including:

- *Statistics in Medicine*
- *Research Synthesis Methods*
- *BMJ*
- *Journal of Clinical Epidemiology*

The analysis follows **PRISMA-NMA** reporting guidelines (Hutton et al., 2015) and implements methods from recent methodological papers.

## Load Package

```{r load-package, eval=FALSE}
library(cnma)
library(dplyr)
library(ggplot2)
```

## Example Dataset

We'll use simulated data representing a network of treatments for a clinical condition. In practice, you would load your own systematic review data.

```{r simulate-data, eval=FALSE}
# Generate simulated data with 40 studies
set.seed(123)
data <- simulate_cnma_data(n_studies = 40, seed = 123)

# Examine data structure
head(data)
dim(data)

# Check included treatments
unique(c(data$treat1, data$treat2))
```

## Step 1: Data Validation and Cleaning

Always validate your data before analysis:

```{r validate-data, eval=FALSE}
# Clean data (removes invalid rows)
data_clean <- cnma_clean_data(data)

# Validate structure and values
validate_cnma_input(data_clean)

# Network characteristics
network_summary <- network_characteristics_summary(data_clean)
print(network_summary)
```

## Step 2: Configure Analysis

Configure analysis parameters following your research question:

```{r configure, eval=FALSE}
config <- setup_cnma(
  sm = "HR",                    # Hazard ratio
  use_bayesian = FALSE,         # Frequentist approach
  run_metareg = FALSE,          # No meta-regression
  export_plots = FALSE,         # Don't save plots
  report_html = FALSE,          # Generate report separately
  seed = 123                    # Reproducibility
)

print(config)
```

## Step 3: Run Main Analysis

```{r main-analysis, eval=FALSE}
# Run network meta-analysis
results <- run_cnma_analysis(
  data = data_clean,
  ref_treatment = "Placebo",    # Specify reference
  config = config
)

# View results
print(results)
summary(results)

# Access underlying netmeta object
nma <- results$results$main_nma
```

## Step 4: Assess Key Assumptions

### 4.1 Transitivity Assessment

Evaluate whether studies are similar enough across comparisons:

```{r transitivity, eval=FALSE}
# Assess transitivity by comparing study characteristics
transitivity <- assess_transitivity(
  data = data_clean,
  variables = c("age_mean", "female_pct", "bmi_mean", "year"),
  by_comparison = TRUE
)

print(transitivity)
```

**Interpretation**: Large differences in study characteristics across comparisons may indicate violation of the transitivity assumption.

### 4.2 Inconsistency Assessment

Comprehensive evaluation of network consistency:

```{r inconsistency, eval=FALSE}
# Multiple methods for inconsistency assessment
inconsistency <- assess_inconsistency(
  nma,
  methods = c("global", "local", "design")
)

print(inconsistency)
```

**Key metrics:**

- **Global**: Overall heterogeneity/inconsistency (Cochran's Q, I², Tau²)
- **Local**: Node-splitting to identify specific inconsistent comparisons
- **Design**: Design-by-treatment interaction test

## Step 5: Treatment Rankings

Calculate treatment rankings using P-scores (frequentist analogue of SUCRA):

```{r rankings, eval=FALSE}
# Calculate P-scores
rankings <- calculate_rankings(
  nma,
  small_values = "undesirable"  # Smaller effects are worse (e.g., for HRs)
)

print(rankings)
```

**P-score interpretation:**
- 1.0 = Certainly the best treatment
- 0.5 = Average performance
- 0.0 = Certainly the worst treatment

## Step 6: Pairwise Comparisons (League Table)

Generate a league table showing all pairwise treatment comparisons:

```{r league-table, eval=FALSE}
# Create league table
league <- create_league_table(nma, digits = 2)
print(league)
```

This format is standard in *BMJ*, *Lancet*, and *JAMA* for presenting NMA results.

## Step 7: Prediction Intervals

Calculate prediction intervals accounting for heterogeneity:

```{r prediction-intervals, eval=FALSE}
# Prediction intervals
pred_intervals <- calculate_prediction_intervals(nma, level = 0.95)
print(pred_intervals)
```

**Interpretation**: Prediction intervals show the expected range of true effects in *future studies*, wider than confidence intervals due to between-study heterogeneity.

Reference: IntHout et al. (2016), *BMJ Open*

## Step 8: Publication-Quality Visualizations

### 8.1 Network Plot

```{r network-plot, eval=FALSE}
# Network geometry
plot_network(nma)
```

### 8.2 Forest Plot with Prediction Intervals

```{r forest-plot, eval=FALSE}
# Forest plot comparing all treatments to reference
plot_forest(
  nma,
  reference = "Placebo",
  prediction_interval = TRUE,
  sortvar = TRUE
)
```

### 8.3 Treatment Ranking Plots

```{r ranking-plots, eval=FALSE}
# Rankogram
plot_rankings(nma, type = "rankogram")

# Cumulative ranking plot
plot_rankings(nma, type = "cumulative")
```

### 8.4 Funnel Plot for Publication Bias

```{r funnel-plot, eval=FALSE}
# Comparison-adjusted funnel plot
plot_funnel(nma)
```

### 8.5 Net Heat Plot for Inconsistency

```{r netheat-plot, eval=FALSE}
# Net heat plot (if sufficient network structure)
plot_netheat(nma)
```

### 8.6 Generate All Plots Automatically

```{r all-plots, eval=FALSE}
# Create all publication plots and save to directory
create_publication_plots(
  nma,
  output_dir = "nma_figures",
  reference = "Placebo",
  device = "both",  # PNG and PDF
  width = 10,
  height = 8
)
```

## Step 9: Sensitivity Analyses

### 9.1 Leave-One-Out Analysis

```{r leave-one-out, eval=FALSE}
# Identify influential studies
loo_results <- leave_one_out_analysis(
  data = data_clean,
  config = config,
  progress = TRUE
)

print(loo_results)
```

**Interpretation**: Studies with >10% change in Tau are considered influential.

### 9.2 Publication Bias Assessment

```{r pub-bias, eval=FALSE}
# Comprehensive publication bias assessment
pub_bias <- assess_publication_bias(
  nma,
  methods = c("visual", "comparison", "egger")
)

print(pub_bias)
```

### 9.3 Model Comparison

```{r model-comparison, eval=FALSE}
# Compare fixed vs random effects models
model_comp <- compare_models(data_clean, config)
print(model_comp)
```

## Step 10: PRISMA-NMA Compliance Report

Generate a comprehensive report for your manuscript:

```{r prisma-report, eval=FALSE}
# Generate PRISMA-NMA compliance report
prisma_report <- generate_prisma_report(
  results,
  output_format = "markdown",
  file = "prisma_nma_report.md"
)

# View in console
cat(prisma_report)
```

## Complete Workflow Example

Here's a complete analysis workflow in one script:

```{r complete-workflow, eval=FALSE}
library(cnma)

# 1. Data preparation
data <- simulate_cnma_data(40, seed = 123)
data_clean <- cnma_clean_data(data)
validate_cnma_input(data_clean)

# 2. Network characteristics
network_summary <- network_characteristics_summary(data_clean)
print(network_summary)

# 3. Configure and run analysis
config <- setup_cnma(sm = "HR", use_bayesian = FALSE, seed = 123)
results <- run_cnma_analysis(data_clean, ref_treatment = "Placebo", config = config)
nma <- results$results$main_nma

# 4. Assess assumptions
transitivity <- assess_transitivity(data_clean,
                                   variables = c("age_mean", "female_pct"))
inconsistency <- assess_inconsistency(nma, methods = c("global", "local", "design"))

# 5. Treatment rankings
rankings <- calculate_rankings(nma)
print(rankings)

# 6. Pairwise comparisons
league <- create_league_table(nma)
print(league)

# 7. Prediction intervals
pred_int <- calculate_prediction_intervals(nma)
print(pred_int)

# 8. Visualizations
create_publication_plots(nma, output_dir = "figures", reference = "Placebo")

# 9. Sensitivity analyses
loo <- leave_one_out_analysis(data_clean, config)
pub_bias <- assess_publication_bias(nma)
model_comp <- compare_models(data_clean, config)

# 10. PRISMA-NMA report
prisma_report <- generate_prisma_report(results, "markdown", "prisma_report.md")

cat("\n✓ Analysis complete! All outputs ready for publication.\n")
```

## Reporting Checklist

When writing your manuscript, ensure you include:

### Required Elements (PRISMA-NMA)

- [ ] Network diagram (use `plot_network()`)
- [ ] Description of network geometry
- [ ] Transitivity assessment (use `assess_transitivity()`)
- [ ] Inconsistency evaluation (use `assess_inconsistency()`)
- [ ] Forest plot with prediction intervals (use `plot_forest()`)
- [ ] League table (use `create_league_table()`)
- [ ] Treatment rankings (use `calculate_rankings()`)
- [ ] Funnel plot for publication bias (use `plot_funnel()`)
- [ ] Sensitivity analyses (use `leave_one_out_analysis()`)
- [ ] GRADE assessment of evidence quality

### Recommended Statistical Details

- Heterogeneity estimates (Tau², I²)
- Between-study variance
- Model assumptions and violations
- Software and version used
- Analysis reproducibility (seed, code availability)

## Key References

1. **PRISMA-NMA Guidelines**:
   Hutton B, et al. (2015). The PRISMA extension statement for reporting of systematic reviews incorporating network meta-analyses. *Annals of Internal Medicine*, 162(11):777-84.

2. **P-scores and Rankings**:
   Rücker G, Schwarzer G (2015). Ranking treatments in frequentist network meta-analysis works without resampling methods. *BMC Medical Research Methodology*, 15:58.

3. **Inconsistency Assessment**:
   Dias S, et al. (2010). Checking consistency in mixed treatment comparison meta-analysis. *Statistics in Medicine*, 29(7-8):932-44.

4. **Net Heat Plot**:
   Krahn U, et al. (2013). A graphical tool for locating inconsistency in network meta-analyses. *BMC Medical Research Methodology*, 13:35.

5. **Prediction Intervals**:
   IntHout J, et al. (2016). Plea for routinely presenting prediction intervals in meta-analysis. *BMJ Open*, 6(7):e010247.

6. **Publication Bias in NMA**:
   Chaimani A, Salanti G (2012). Using network meta-analysis to evaluate the existence of small-study effects. *Research Synthesis Methods*, 3(2):161-76.

## Session Info

Always report your session information for reproducibility:

```{r session-info, eval=FALSE}
sessionInfo()
```

## Conclusion

The **cnma** package provides a comprehensive, journal-ready toolkit for network meta-analysis that:

- ✓ Follows PRISMA-NMA reporting guidelines
- ✓ Implements state-of-the-art statistical methods
- ✓ Generates publication-quality visualizations
- ✓ Provides comprehensive sensitivity analyses
- ✓ Ensures reproducible research

For more information, visit: <https://github.com/mahmood726-cyber/HFN786>

---

**Citation**: When using cnma in your research, please cite:

CNMA Development Team (2025). cnma: Comprehensive Network Meta-Analysis with PRISMA-NMA Compliance. R package version 1.1.0. https://github.com/mahmood726-cyber/HFN786
